---
title: "Morse Transcriber"
output: html_notebook
---

This workbook converts an audio recording of The Last of Us Morse Cricket to text.

This code is based almost entirely on Mark Patterson's work, all I did was modify a few settings and remove some unneeded translation stuff.

https://www.r-bloggers.com/morse-code-converter/

Steps:
Record video of the Morse Cricket.  For best results, turn all audio volumes except effects down to 0, turn off your flashlight (annoying flickering), and stand facing the corner so that the right channel has the strongest audio.

Rip an MP3 from the recording.  I broadcast my videos to YouTube, and then used https://www.onlinevideoconverter.com/mp3-converter to just download an MP3.

Convert to WAV format, I used ffmpeg.

Remove background noise.  I used Audacity and applied a 4200-5200hz bandpass filter.

Run the following blocks of the script as needed.

Good luck!

```{r}
# install packages if you don't have them

install.packages("tuneR")
install.packages("RCurl")
install.packages("seewave")
```

```{r}
#### initialize ####
 
# clear workspace:
rm(list = ls())
 
## loading libraries:
library(tuneR)
library(RCurl)
library(seewave)

# helper function - finds the variance of a 100-sample slice of the audio:
var_find = function(vec, t, s){
  var.out = var(vec[(t-s):(t+s)])
  return(var.out)}
```

```{r}
# load up a file

infile = "~/repos/morsecricket/data/60m.filtered"
sound.file = readWave(paste(infile, ".wav", sep = ""))

#sound.filtered = ffilter(sound.file, from = 4500, to = 5000, bandpass = TRUE, rescale = TRUE, output = "Wave")
#writeWave(filtered, "~/repos/morsecricket/20sec.filtered.wav")
#sound.file = readWave("~/repos/morsecricket/20sec.filtered.wav")
```

```{r}
# read data into a dataframe
df = data.frame(indx = 1:length(sound.file@right), vec = sound.file@right)

# points to sample:
sample.points = seq(from = 100, by = 100, to = length(df$vec))

# applying the variance finder at the sampled points:
variance = data.frame(var = sapply(sample.points, 
                                  function(x){var_find(vec = df$vec,
                                                       t = x,
                                                       s = 50)}))
```

```{r}
# decide which points are 'on'
on = as.numeric(variance > 100)

# special for the cricket.  the chirps are actually a bunch of tiny chirps strung together.
# fill the small holes between them with 1s, then eliminate small values between big chirps.
on = inverse.rle(within.list(rle(on), values[lengths<20] <- 1))
on = inverse.rle(within.list(rle(on), values[lengths<20] <- 0))

tiny.df = data.frame(on = on)
tiny.df$indx = 1:nrow(tiny.df)

# create a vector of changes in on:
raw.vec = diff(tiny.df$on)
```

```{r}  
# figure out beeps

# create indices for change instances -- these will be 1 and -1
beep.start.vals = which(raw.vec == 1)
beep.stop.vals = which(raw.vec == -1)

# trim the last start value if its longer
beep.start.vals <- beep.start.vals[1:length(beep.stop.vals)]

# converting indices to durations:
beep.durs = beep.stop.vals - beep.start.vals

  ## note: for some files, there seems to be a few 
## few beep durs that are only 1; for now, hard coding these out:
beep.durs = beep.durs[beep.durs>10]

# short chirps are about 50ish, long chirps are 170+.  Just use 100 as a separator
beep.sep = 100

# recoding beep durations to long and short:
beep.durs.let = beep.durs
beep.durs.let[beep.durs > beep.sep] = "-"
beep.durs.let[beep.durs < beep.sep] = "."

# interestingly, chirps come in 180ish, 200ish, and 240ish durations.  Meaningful?
#plot(beep.durs)

write(paste(unlist(beep.durs.let), collapse = ""), paste(infile, ".txt", sep = ""))
``` 
